Demo for a Quality of Service networking with Software Defined Network.

### 2017-03-15 report/paper.tex
final correction. everything should be fine
images sistemate manca da unire alle correzioni di andrea faccio dopo

\begin{abstract}
Quality of Service (QoS) in existing network architectures is an ongoing problem.
A lot of researchers and industries tried to solve this problem and found different solutions but some of them are too
expensive other failed or were not implemented.
Software Defined Networking (SDN) paradigm has emerged in response to traditional networking architecture's limitations.
The main advantage of SDN is a centralized global network view, which pushed researchers to keep on improving QoS exploiting it.

In this report we will focus on two different aspects, the problems we faced during our attempt to create a based SDN application able to
support a QoS service and the results we achieved.

OpenFlow standard and python POX controller, which is largely spread, were used as instruments to make this experience possible,
despite the limit of using a new type of technology still in the research phase.
%corretto
\end{abstract}

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is a way to manage networks that separates the control from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
The SDN Controller relays on the information given by switches and routers via southbound APIs, and acts with what the applications set via northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it is not the only SDN implementation available. 
Centralized, programmable SDN environments can easily adjust to the rapidly changing business needs and can reduce costs, restrict
wasteful provisioning, as well as provide flexibility and innovation for networks.
%corretto
\subsection{OpenFlow}
  During 2006 ,thanks to M. Casado, a new network architecture was developed, called Ethane.
  On this conncept was born, during 2011, the Open Network Foundation, a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 was released on 2009 but the last standardized version v1.5.1 was approved only on march 2015 \cite{ONF_report}.
  %corrects
  %todo spostare l'immagine correttamente
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.58]{images/of.jpg}
 \caption{\emph{OpenFlow}}
 \label{fig:topo}
\end{figure}
  \\
  \textbf{Openflow 1.0 messages}

Openflow defines a number of messages between a controller and different switches; some of the ones we exploited will be shown right after:\cite{openflow}:
\begin{itemize}

\item \emph{Hello} \newline This message is sent both from the controller and the switch who is looking for a controller after it
	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its capabilities. The switch
	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
\item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
		It's useful to check if a port has someone connected to it.
\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of 
statistic the controller wants to know.
		The switch then will answer with the Stats reply, if it supports that type of stat. If not, The switch will send a Bad_Request
		message.
\item \emph{PacketIn} \newline This message is sent from the switch to the controller everytime a packet arrives without matching any rule.
The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
		a flow rule for packets like this using a FlowMod.
\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message,
to some ports.
\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch, allowing traffic to be routed along specific paths. The structure of this message is explained as follows.
\\
\end{itemize}

\textbf{Forwarding table}
\begin{itemize}
 \item \texttt{match fields:} compare packets properties in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
 \item \texttt{priority:} this field is used when the specific packet matches more than one rule. The higher the number, the higher the priority a rule has;
 \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
 \item \texttt{istruction:} a set of actions that the switch add on \emph{Action-set} of the matched packet or some instruction 
 in order to modify the pipeline flux;
 \item \texttt{timeout:} this is the maximum time afther which the rule expire;
 \item \texttt{flag:} with this field we can differenciate similar voices in order to managed multiple table with similar rules;
 \\
 \\
\end{itemize}

\textbf{Match Constraints}
these are the principal matching method that we can use with OpenFlow v1.0:
\begin{itemize}
 \item \texttt{in_port}: the port of swith the packet is arriving;
 \item \texttt{dl_src/dl_dst}: IEEE 802 MAC source and destination of packet;
 \item \texttt{dl_type}: IEEE 802 EtherType;
 \item \texttt{nw_tos}: IPv4 ToS;
 \item \texttt{nw_proto}: IPv4 Protocol;
 \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination addresses;
 \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination ports;
 \\
 \\
\end{itemize}

\textbf{Action} With OpenFlow v1.0 we can do basic tasks in practice, the most important are for example:
\begin{itemize}
 \item \texttt{Output}: set output port;
 \item \texttt{SetDLSrc/Dst}: modify packet's MAC address;
 \item \texttt{SetNWSrc/Dst}: modify packet's IPv4 address;
 \item \texttt{SetNWTos}: change ToS fields of message;
 \item \texttt{SetTPSrc/Dst}: set the TCP or UDP source and destination port.
\end{itemize}

@@

Another limitation is the no default routing from a source to a destination: when an host
is plugged to a network of SDN switches, it won't know the destination. The only way it can
discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
the network topology, but doesn't know how much traffic each host is doing. For that the administrator need to put a firewall, which it's not needed in SDN networks,
as the controller can set a switch to drop all the packets from/to a destination.

@@

\subsection{Router Board}
\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.30]{images/rboard.png}
  \caption{\emph{Experience scenario}}
  \label{fig:topo}
\end{figure}

@@

When we set up this option \texttt{OFPP_FLOOD} like output port's action in a table's switch row if some spanning_tree algorithm is working
the \texttt{OFPP_FLOOD} is only theoretical in the sense that the algorithm excludes some ports during the flooding option, in order to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is as follows:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

@@

  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/topology.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
  \end{figure}

@@

\subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline
Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in order to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
What we observe is a estimated delay like a few seconds which is not the case, infact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\
\\
The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.26]{images/delay.png}
 \caption{\emph{Delay's graph.}}
 \label{fig:topo}
\end{figure}

		\subsubsection{Link Capacity}

>>>>

\begin{abstract}
Quality of Service (QoS) in existing network architectures is an ongoing problem.
A lot of researchers and industries tried to solve this problem and found different solutions but some of them are too
expensive other failed or were not implemented.
Software Defined Networking (SDN) paradigm has emerged in response to traditional networking architecture's limitations.
The main advantage of SDN is a centralized global network view. This particular feature pushed researchers to keep on improving QoS.

In this report we will focus on two different aspects; all problems we faced during our attempt to create a based SDN application able to
support a QoS service and the results we achieved.

OpenFlow standard and python POX controller, which is largely spread, were used as instruments to make this experience possible,
despite the limit of using a new type of tecnology still in the research phase.
%corretto
\end{abstract}

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is a way to manage networks that separates the control from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it isn’t the only SDN standard. 
Centralized, programmable SDN environments can easily adjust to the rapidly changing business'needs. SDN can reduce costs and restrict
wasteful provisioning, as well as provide flexibility and innovation for networks.
%corretto
\subsection{OpenFlow}
  During 2006 ,thanks to M. Casado, a new network architecture was developed, called Ethane.
  On this conncept was born, during 2011, the Open Network Foundation, a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 was released on 2009 but the last standardized version v1.5.1 was approved only on march 2015 \cite{ONF_report}.
  %corrects
  %todo spostare l'immagine correttamente
\begin{center}
 \centering
 \includegraphics[scale=0.58]{images/of.jpg}
 {Figure 1: \emph{OpenFlow}}
 %\label{fig:topo}
\end{center}
\textbf{Openflow 1.0 messages}

Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited\cite{openflow}:
\begin{itemize}

\item \emph{Hello} \newline This message is sent both from the controller and the switch who is looking for a controller after it
	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its abilities. The switch
	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
\item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
		It's useful to check if a port has someone connected to it.
\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of 
statistic the controller wants to know.
		The switch then will answer with the Stats reply, if it supports that type of stat. If not, The switch will send a Bad_Request
		message.
\item \emph{PacketIn} \newline This message is sent from the switch to the controller everytime a packet arrives without matching any rule.
The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
		a flow rule for packets like this using a FlowMod.
\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message,
to some ports.
\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch.This is the message that allows routing.
\\
\\
\end{itemize}

\textbf{Forwarding table}
\begin{itemize}
 \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
 \item \texttt{priority:} this field is used when there are rows with more than one correspondence;
 \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
 \item \texttt{istruction:} a set of actions that the switch add on \emph{Action-set} of the matched packet or some instruction 
 in order to modify the pipeline flux;
 \item \texttt{timeout:} this is the maximum time afther which the rule expire;
 \item \texttt{flag:} with this field we can differenciate similar voices in order to managed multiple table with similar rules;
 \\
 \\
\end{itemize}

\textbf{Match Constraints}
these are the principal matching method that we can use with OpenFlow v1.0:
\begin{itemize}
 \item \texttt{in_port}: the port of swith;
 \item \texttt{dl_src/dl_dst}: 802.3 MAC source and destination of packet;
 \item \texttt{dl_type}: 802.3 EtherType;
 \item \texttt{nw_tos}: IPv4 ToS;
 \item \texttt{nw_proto}: IPv4 Protocol;
 \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination addresses;
 \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination ports;
 \\
 \\
\end{itemize}

\textbf{Action} With OpenFlow v1.0 we can do basic tasks in practice, the most important are for example:
\begin{itemize}
 \item \texttt{Output}: set output port;
 \item \texttt{SetDLSrc/Dst}: modify packet's MAC address;
 \item \texttt{SetNWSrc/Dst}: modify packet's IPv4 address;
 \item \texttt{SetNWTos}: change ToS fields of message;
 \item \texttt{SetTPSrc/Dst}
\end{itemize}

@@

Another limitation is the no default routing from a source to a destination: when an host
is plugged to a network of SDN switches, it won't know the destination. The only way it can
discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.

@@

\subsection{Router Board}
\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.30]{images/rboard.png}

  \label{fig:topo}
\end{figure}
\begin{center}
 {Figure 2: \emph{Experience scenario}}
\end{center}

@@

When we set up this option \texttt{OFPP_FLOOD} like output port's action in a table's switch row if some spanning_tree algorithm is working
the \texttt{OFPP_FLOOD} is only theoretical in the sense that the algorithm excludes some ports during the flooding option, in order to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is as follows:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 {Figure 3: \emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

@@

  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/topology.png}
 {Figure 4: \emph{Experience scenario}}
 \label{fig:topo}
  \end{figure}

@@

\subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{center}
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}

 \label{fig:topo}
\end{center}

\begin{center}
 {Figure 5: \emph{RTT between controller and OF switch}}
\end{center}

If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline
Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in order to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \label{fig:topo}
\end{figure}

\begin{center}
  {Figure 6:\emph{Pck from controller to itself following a particular way.}}
\end{center}

At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
What we observe is a estimated delay like a few seconds which is not the case, infact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\
\\
The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
\begin{center}
 \centering
 \includegraphics[scale=0.26]{images/delay.png}
 {Figure 7:\emph{ Delay's graph;}}
 \label{fig:topo}
\end{center}

%\end{figure}
\subsubsection{Link Capacity}
